Mom, Let Me Play More Computer Games: They Improve My Mental Rotation Skills
Isabella D. Cherney.


\section{Introduction}
Visuospatial and mathematical skills may serve as mediators of 
gender-based mathematics differences. Both these skills are critical for
mastering material in the physical sciences. 
Massed or distributed practice? 
Individual differences, affect outcomes?
Let's test this thing! With 2 and 3 D games.

Cognitive Gender differences:
Women perform better on verbalt tests
Men have better visuospatial capabilities (Mental rotation tests (Vandeberg 
and Kuses VMRT))
More striking at both the lower and upper extremes of intellectual ability
Q: WHY?!
A: from early on different kind of behavior 

Training and practice effects
after training less or no differences and better
results
Training effects are larger than gender differences
Video games are practice...  (and better practice for females)

Massed/distributed practice
Unclear for massed/distributed practice! 

Individual differences
few research has examined this
Anxiety
Super math skills
experiences in sports/gamesA

Present study
2/3d and delivery (massed vs distributed practice) and individual differences
(prioor experience, anxiety, mathematical performance)
Spatial task: VMRT (vandeberg Mental rotation test: 3d) and CRT (card rotation test: 2d) 

Cognitive gender difference hypotheses
significant correlation between the rotation tests
negative relation between anxiety and performance (DUH?)

practice hypotheses
practice is beneficial for all genders
Females might benefit more because they don't have previous practice

massed vs distributed hypothesis
distributed ftw

individual differencens hypotheses
lower levels of anxiety, higher levels of computer game/sports expereince 
and higer levels of mathemetical performance were expected to be predictive
of overall positive performance gains

/section{method}

participants
61, 30 men and 31 women
aged between 17 and 23 (mean age 19.1, SD = 1.4)
random: 3d (20 folks) and 2d (21 folks) and control (20 folks)
36 over 2 weeks (distributed)
25 1 week (massed)

Material
CRT: an abstract 2d figure, rotated (same) or mirror image (different)
8 examples and 20 targets
VMRT match a 3d figure to 2 possible other figures.
2 ten items sections
Background questionnaire
Math test 
STAI (state- and trait-anxiety inventory), 
20 questions, "how do you feel?" 
Games:
Antz Extreme racing for 2d 
tetrus (tetris) for 2d
control condition: Game world of puzzles (it's a book!)

Procedures
Distributed: 3 1 hour sessions in 2 weeks 
Massed: the sessions within 3 days

\section{Results}
* Noone knew antz (duh) and an equal amount of males/femals already knew tetris
* Cognitive gender difference analyses 
 - VMRT and CRT scores correlated significantly
 - anxiety levels and performance correlated negatively
 - more difficult test means more difference males/femals
 - VMRT showed gender differences and CRT did not (!)
* practice effect analysis
 - women showed greater gains
 - the control group's performance increased significantly
 
* massed vs distributed analysis
 - more gain when massed
* individual differences
 - math, gender and practice conditions could predict 33% of the variance
 - Gender and 3d training conditions were significant contributors. 
 - Gender and type of practice were not predictive of the performance

\section{discussion}
short practice (3 hours) computer game play improve performance on 
MRT. Women more than men. Massed training was bettor. 
Individuals better at math, gender and type of training condition perform 
better. 

 Seems to fit with previous studies and also hypothesis (which is how what 
 they based the hypothesis on). 

No significant gender differences for CRT. 

 Perhaps it was tooo easy 
 (otherwise corresponded with the VMRT).

Gains in women were higher than those of men but the final result was 
 still lower than mens result.


The 3d game antz was awesome for improving on the VMRT.   
Short term is good enough to increase scores on the short term.
women have higher anxiety treat
men have more experience in gaming

 cannot show people gaming more because they are better at spatial thingies

math is important

Limitations

 antz is not really a navigational game ....

short duration of learning
Could retaking the test always improve performance?

Future study
longer learning
better games
various types of training































